\paperTitle

### 2014-05-09 doc/icml2014/directed.tex
Commented out assumptions proofs and touched up piecewise section

  Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
  Consider the following transformation.
  Replace $x_v$ with a new hidden variable \hp\ with the same
  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
  (\figureref{reduction}). 
  The domain of $\hp$ is the same as $x_v$\footnote{
      Though we have assumed that all hidden variables share the
      same domain, $[k]$, our work generalizes to hidden domains of
      different sizes, provided \assumptionref{full-rank-plus} is satisfied.
      },
    and $\mOpp{v}{\textrm{new}} = I$.
  Then, there is a one-to-one correspondence between every value of
  $\hp$ and $x_v$. Consequently, for any clique $\sC \contains \hp$, the
  parameters in the original graphical model can be obtained by
  substituting $\hp$ with $x_v$.

@@

To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full
  column rank and that $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
  $x_v$, marginalizing out the rest of the graph.
To make this property more interpretable, let us assume the following
  property which we prove to be sufficient in
  \appendixref{assumption-proof}:
\begin{assumption} 
\label{asm:full-rank-plus}
The marginal distribution of every latent variable has full support:
  $\forall{h_i \in H},~ \mPi{i} \succ 0$.
For every clique $\sC \in \sG$ (including ones with observed variables),
  and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
  $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
\end{assumption}
In our directed grid example (\sectionref{directedExample}), for the
  clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
  \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
  $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
  rank $k$.

@@

From \citet{anandkumar13tensor} given \assumptionref{full-rank-plus}, we
  have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of
  $n^{-\frac12}$ with a constant that depends polynomially on the $k$-th
  singular value of $\mOpp{v}{i}$:

\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
  \label{thm:sample-complexity}
  Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
  $h_1$. With probability at least $1 - \delta$,
\begin{align*}
  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
    &\le  \\
    &
      O\left(
      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min} } 
      {\sqrt{n}(\sigma_{\min}(\mOpp{1}{1}) \sigma_{\min}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
\end{align*}

If $\sigma_{\min}(\Pr(c \given \Pa(h_c))) \le \sigma$ for every clique
  $\sC \in \sG$ and hidden variable $h_c \in \sC$,
\begin{align*}
  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
    &\le 
      O\left( 
      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
      {\sqrt{n} \sigma^{5t}} \right) \epsilon,
\end{align*}
where $t$ is the length of the shortest topological ordering from $h_1$
to the unique parent of $x_1$, $h_t$.
\end{theorem}

\theoremref{sample-complexity}, proved in \appendixref{assumption-proof}
  says that the number of samples required to estimate the view
  $\mOpp{v}{i}$ grows exponentially in the {\em distance} between $x_v$
  and $h_i$. 
Intuitively, this is expected because the influence from $h_i$ to $x_v$
  gets diluted for every latent variable it must factor through.

\todo{Present sample complexity bounds for $Z_\sC$?}

>>>>

  Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
  Consider the following transformation.
  Replace $x_v$ with a new hidden variable \hp\ with the same
  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
  (\figureref{reduction}). 
  The domain of $\hp$ is the same as $x_v$\footnote{
      Though we have assumed that all hidden variables share the
      same domain, $[k]$, our work generalizes to hidden domains of
      different sizes, provided \assumptionref{full-rank} is satisfied.
      },
    and $\mOpp{v}{\textrm{new}} = I$.
  Then, there is a one-to-one correspondence between every value of
  $\hp$ and $x_v$. Consequently, for any clique $\sC \contains \hp$, the
  parameters in the original graphical model can be obtained by
  substituting $\hp$ with $x_v$.

@@

To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full
  column rank and that $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
%   $x_v$, marginalizing out the rest of the graph.
% To make this property more interpretable, let us assume the following
%   property which we prove to be sufficient in
%   \appendixref{assumption-proof}:
% \begin{assumption} 
% \label{asm:full-rank-plus}
% The marginal distribution of every latent variable has full support:
%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
% For every clique $\sC \in \sG$ (including ones with observed variables),
%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
% \end{assumption}
% In our directed grid example (\sectionref{directedExample}), for the
%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
%   rank $k$.

@@

From \citet{anandkumar13tensor} given \assumptionref{full-rank}, we
  have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of
  $n^{-\frac12}$ with a constant that depends polynomially on the $k$-th
  singular value of $\mOpp{v}{i}$:
\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
  \label{thm:sample-complexity}
  Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
  $h_1$. With probability at least $1 - \delta$,
\begin{align*}
  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
    &\le  \\
    &
      O\left(
      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min} } 
      {\sqrt{n}(\sigma_{\min}(\mOpp{1}{1}) \sigma_{\min}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
\end{align*}

%If $\sigma_{\min}(\Pr(c \given \Pa(h_c))) \le \sigma$ for every clique
%  $\sC \in \sG$ and hidden variable $h_c \in \sC$,
%\begin{align*}
%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
%    &\le 
%      O\left( 
%      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
%      {\sqrt{n} \sigma^{5t}} \right) \epsilon,
%\end{align*}
%where $t$ is the length of the shortest topological ordering from $h_1$
%to the unique parent of $x_1$, $h_t$.
\end{theorem}

In general, $\sigma_{\min}(\mOpp{v}{i})$ can depend exponentially on the
  path length between the hidden variable $h_i$ and $x_v$.
Intuitively, this is expected because the influence from $h_i$ to $x_v$
  gets diluted for every latent variable it must factor through.

%\theoremref{sample-complexity}, proved in \appendixref{assumption-proof}
%  says that the number of samples required to estimate the view
%  $\mOpp{v}{i}$ grows exponentially in the {\em distance} between $x_v$
%  and $h_i$.

%\todo{Present sample complexity bounds for $Z_\sC$?}
